
\documentclass[11pt,a4paper,hidelinks,fleqn]{article}            % Article 12pt font for a4 paper while hiding links
\usepackage[margin=1in]{geometry}                          % Required to adjust margins
\input{styleAndCommands}
\title{\vspace{-5ex}Exam, FE5116 (2014/2015, Semester 2)\vspace{-7ex}}
\date{}
\begin{document}
\maketitle
\subsection*{Question 1 [15 marks]} 
We want to model the process of stock price using the following SDE:
\begin{align*}
\frac{dS(t)}{S(t)} = \mu(t) dt + \sigma(t) dW_t
\end{align*}
and would like to use $X(t) = \ln(S(t))$ as the state variable.

a) What is the SDE for $X(t)$? [8 marks]
\vspace{4.5cm}

b) Given a standard normal random variable $z \sim N(0, 1)$ 
and time step $\Delta t$,
write down the equation that evolves $X$ from 
time $t$ to $t+\Delta t$ using Euler scheme, i.e., 
the function $f$ that $X_{t+\Delta t} = f(X_t, \Delta(t), z)$. [7 marks]

\newpage

\subsection*{Question 2 [15 marks]} 
Compute the integral $\int_0^2 (-x^2+2x+5) dx$ using composite integration formula with 2 intervals of the following quadrature:

a) Mid-point rule [5 marks]
\vspace{2cm}

b) Trapezoid rule [5 marks]
\vspace{2cm}

c) Simpson rule (using 3 points for each interval, 5 points in total) [5 marks]
\vspace{2cm}

\subsection*{Question 3 [15 marks]} 
a) Set up a Newton's method to find the intersection of the curve $y=x^3$ and the line $y = - x + 4$. [10 marks]
\vspace{4cm}

b) Start with $x_0 = 0$, what are $x_1$ and $x_2$ after first and second iterations of the Newton's method? [5 marks]

\newpage
\subsection*{Question 4 [20 marks]}
The price of a dividend free stock S follows a geometric Brownian motion with volatility 20\%.
The risk free continuous compounding interest rate is 5\%. The stock price today is 100 dollars.

a) Set up a two step JR (Jarrow, Rudd) binomial tree to price options with 2 year maturity using the stock price as state variable.
You should list and compute the parameters of the tree, and draw out the nodes of the tree with their associated stock price. [10 marks]
\vspace{8.5cm}

b) Price a 2 year European call option (strike = 95 dollars) on the tree you have set up in a). List the option price at each node of the tree.  [10 marks]
\newpage

\subsection*{Question 5 [20 marks]}
We would like to price an Asian option on a stock $S(t)$ using two moment matching method.
The option's payoff is based on the average of the stock price observed at a sequence of observation times: $A = \frac{1}{n}\sum S(t_i)$.
Two moment matching approximate the distribution of $A$ using a log normal distribution whose first and second moments matches the first and second moments of $A$.
The interest rate is 0\%. The current stock price is $S_0$.
The stock price process follows the Black-Scholes SDE
\begin{align*}
\frac{d(S(t))}{S(t)} = \sigma dW_t
\end{align*}

a) Derive the formula for $E[A]$ (the first moment of A)? [5 marks]
\vspace{3cm}

b) Derive the formula for $E[A^2]$ (the second moment of A)? [15 marks]
\newpage


\subsection*{Question 6 [15 marks]}
Our underlying asset $X_t$ follows the Ornstein-Uhlenbeck process
\begin{align*}
d X_t = \theta(\mu - X_t) dt + \sigma dW_t
\end{align*}
where $\theta$, $\mu$ and $\sigma$ are all positive constants. The interest rate $r$ is constant as well.
Derive the PDE that the price of a derivative on $X_t$ (denoted as $V_t$) should satisfy. 


\end{document}
